---
title: "project1"
author: "Joo Kai Tay (22489437)"
date: "2023-08-07"
output: html_document
---

### Introduction
The data set analyzed can be obtained from the Kaggle platform.
https://www.kaggle.com/datasets/nelgiriyewithana/global-youtube-statistics-2023

This dataset contains information about 995 of the most subscribed YouTube channels in the world. The dataset includes variables such as suscriber count, video views, upload frequency, country of origin and earnings. 

### Data loading, overview and set up

Load libraries
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(ggthemes)
library(numform)
library(treemapify)
library(timeDate)
library(lubridate)
library(dplyr)
library(reshape2)
library(ca)
```

Setting up a plotting theme so that all charts look coherent
```{r}
cits4009_theme <- theme_few() + # Theme based on S. Few's "Practical Rules for Using Color in Charts"
                  theme(plot.title = element_text(color = "darkred")) +
                  theme(strip.text.x = element_text(size = 14, colour = "#202020")) +
                  theme(plot.margin=margin(10,30,10,30))
```